|

### 2021-11-10 article/article.Rmd
format figures and add affiliations

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author:
date: \today
abstract: "How does legislature size impact public service provision? Despite the importance of institutional design for democratic governance, the effect of particular legislative features on citizen welfare remains little understood. In this paper, we use a formal model to show that increasing legislature size improves public goods provision. We argue that changes in bargaining costs depend on whether additional legislators share the Executive's party affiliation: More opposition members reduce the Executive's public goods provision, while more government-aligned members increase it. We test this theory by exploiting sharp discontinuities in city council size in Brazil. We show that an additional city councilor has a 91% chance of belonging to the mayoral coalition, and this significantly improves primary school enrollment and infant mortality measures. To explore possible mechanisms, we surveyed former councilors and analyzed 346,553 bills proposed between 2005 and 2008. This paper has implications for the design of representative institutions."
abstractspacing: double
keywords: Legislature Size; Legislative Bargaining; Local Public Goods Provision; Regression Discontinuity Design; Brazil
jelcodes: H41; H80; I31; I38
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

@@

Third, our article also contributes to a growing literature on municipal partisan politics. Most papers in this area estimate the effects of the mayor's party on political outcomes. For instance, \citet{gerber2011mayors} and \citet{de2016mayoral} find differences in governance when comparing Democratic and Republican mayors. Specifically about Brazil, \citet{gouvea2021partisanship} shows that social policy expenditure increase with left-wing mayors. \citet{frey2020larger}, using discontinuities in city council size after the 2009 constitutional amendment that reverted the thresholds we study here, shows that changes in council size lower the support for candidates backed by the municipal mayors. Our paper reinforces the view that partisanship matters locally, even in countries with weak ideological linkage.

Finally, this paper contributes to a growing literature on the causes and effects of legislature sizes. Most papers on legislature size study the trade-off between representation and efficiency \citet{Stigler1976, muzzio1989size, frederick2008people, rush2013conclusion, jacobs2015explaining}. However, a few new papers explore the effects of legislature size on lawmaking and partisanship \citep{rush2013conclusion, kirkland2014chamber}. For instance, \citet{kirkland2014chamber} shows that larger legislatures make it harder for politicians to know each other and collaborate, making them rely more heavily on party cues and committee structures. In considering the causes of changes in legislature sizes, \citet{jacobs2015explaining} show that population shifts and economic crises cause enlargement and shrinkage, respectively. \citet{marland2019fewer} questions the economic crises mechanism, showing that political elites shrink legislatures mostly to signal to the public the priority of their austerity agenda. In this paper, we show that legislature size affects welfare by changing the partisan composition of legislatures.

@@

The design of institutions is crucial for long term welfare \citep{north1989constitutions, buenodemesquita2005logic, acemoglu2012nations} and democratic consolidation \citep{lijphart2018institutional, gandhi2019institutional}. One particular problem that each polity faces is determination of the size of its legislative bodies. The literature is extensive, and modern attempts to provide a solution to the legislature size problem started as early as in the eighteenth century, with the Federalists versus Anti-Federalists debates \citep{Stigler1976, muzzio1989size}.

The legislature size literature highlights that the number of legislators causes tension between representation and efficiency \citep{taagepera1972size, Stigler1976, muzzio1989size, frederick2008people, rush2013conclusion, jacobs2015explaining}. Larger legislatures improve representation as the election thresholds decrease and the ratio of population-per-politician decreases. This facilitates the electoral success of ethnic and political minorities \citep{allenrepresentation}. Conversely, the improvements in representation are countered by the decline in efficiency in larger legislatures. The enlarged size makes it harder for members to coordinate and collaborate without partisanship and legislative structures \citep{rush2013conclusion, kirkland2014chamber}. It also increases the potential number of veto players, decreasing the policymaking efficiency \citep{tsebelis2002veto}. Balancing these two incentives is crucial for the optimal legislative composition \citep{taagepera1972size, auriol2012optimal}.

However, the legislature size scholarship has overlooked two questions: first, does legislature size matter in contexts where the legislative bodies are not the pivotal decision-maker? And second, what is the ultimate effect of legislature size on welfare? On the one hand, legislature size's effects on welfare represent a significant gap in the political economy literature. It is likely generated by the predominance of the "law of 1/n" scholarship, which emphasized larger legislatures' overspending consequence. Since the seminal paper by \citet{Weingastetal1981}, 30 papers have systematically tested the effects of legislature size on government spending \citep{freire2021legislature}. Of these papers, only \citet{lewis2019legislature} investigated its effects on welfare, showing that larger legislatures have a negative welfare effect.\footnote{\citet{koppel2004public} also shows, using a repeated game where the proposal prerogative rotates in a legislature, that larger legislatures increase the chances of non-cooperative outcomes, lowering public goods provision.} Moreover, \citet{frey2020larger} finds, for the Brazilian case after the 2009 legislature size and salaries constitutional amendment, a null effect in service provision from 2012 to 2016. These findings suggest that the legislature size and welfare nexus need a dedicated theoretical and empirical assessment.

On the other hand, albeit the division of power, in most mayor-council (or presidential) systems, the Executive dominates the service provision. For instance, while the New York Times cited the New York City Mayor Bill De Blasio 2,084 times in 2021, it mentioned the 49 seating city councilors 440 times in the same period. To avoid the COVID crisis when the Executive was in the spotlight, we expanded the search for the past five years. In this new search, Bill De Blasio was mentioned 8,978 while the 49 seating city councilors were mentioned 2,046 times.\footnote{These figures were computed using the Dow Jones Factiva Database in November 2021.} The mayor was cited, on average, over four times more than all the city councilors combined. This Executive dominance has implications for policymaking, as coalitions and affinity with the Executive are going to be crucial for service provision \citep{buenodemesquita2005logic}.

@@

The solution concept we use in this model is the subgame perfect Nash equilibrium. This game requires that the strategies follow a Nash equilibrium in each subgame. We find the solution using backward induction. In the model with infinitely repeated proposals, any strategy can be shown to be a Nash equilibrium. Thus, a natural assumption is to extend the equilibrium concept to require the equilibrium to be stationary. A stationary subgame perfect equilibrium requires that, at each given point of the game, if a politician accepts an offer at time $k+1$, she should take the same offer at time $k$ \citep{rubinstein1982perfect}. Stationarity gives us a method to find a proposal that would be accepted at any stage of the game. Hence, we can characterize a sequence of offers for each point $k$ in time. For the mayor, the optimal offer is at $k=0$, representing a no delay in policy implementation. Moreover, we assume that the players play no dominated strategies, even when there is insufficient support for the policy.

@@

\begin{table}[!htbp] \centering 
  \caption{Political Effects of Increasing the Number of Legislators} 
  \label{tab:bargain} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
& \multicolumn{4}{c}{\textbf{Panel A: Representation, Competition, and Legislation Approval}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Num. Female}} & 
 \multicolumn{1}{c}{\small{Num. Non-white}} &
  \multicolumn{1}{c}{\small{Candidates}} & 
  \multicolumn{1}{c}{\small{Prop. Laws}} \\ 
   & \multicolumn{1}{c}{\small{councillors}} & 
 \multicolumn{1}{c}{\small{councillors}} &
  \multicolumn{1}{c}{\small{Per Seat}} & 
  \multicolumn{1}{c}{\small{Approved Council}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & 0.22 & 0.60 & -0.19 & -0.02 \\ 
  & (0.38) & (0.85) & (1.11) & (0.08) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5183} & \multicolumn{1}{c}{239} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{3424} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{158} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{269} \\ 
Eff N Left & \multicolumn{1}{c}{195} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{179} & \multicolumn{1}{c}{227} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{132} & \multicolumn{1}{c}{141} \\ 
BW Loc Poly & 8.623 & 2.958 & 8.144 & 10.976 \\ 
BW Bias & 13.565 & 4.922 & 12.018 & 16.708 \\[1.0ex] 
\midrule 
& \multicolumn{4}{c}{\textbf{Panel B: Partisan Alignment and Bureaucratic Appointments}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(5)}} 
 & \multicolumn{1}{c}{\small{(6)}} & & \\
 & \multicolumn{1}{c}{\small{Mayoral}} & 
 \multicolumn{1}{c}{\small{Num. Politically}} & & \\
 & \multicolumn{1}{c}{\small{Coalition Size}} & 
 \multicolumn{1}{c}{\small{Bureaucrats}} & 
  & \\[1ex] \cline{2-2} \cline{3-3}
 LATE & 0.91$^{*}$ & 104.49$^{*}$ & & \\ 
  & (0.50) & (61.53) & & \\ 
    \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5179} & \multicolumn{1}{c}{15536} & & \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{1027} & & \\ 
Eff N Left & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{372} & & \\ 
Eff N Right & \multicolumn{1}{c}{162} & \multicolumn{1}{c}{343} & & \\ 
BW Loc Poly & 10.025 & 6.242 & & \\ 
BW Bias & 15.889 & 10.243 & & \\ 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

@@

\begin{figure}[htb!]
  \begin{center}
  \caption{Legislation Approval Patterns in the Municipalities Closer to the City Council Size Thresholds}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=6in,angle=0]{../figures/lawsApproved.pdf}}
      \bigskip
      \label{LEGAPPR}
    \end{minipage}
  \end{center}
\end{figure}

@@

\begin{table}[!htbp] \centering 
  \caption{Legislature Size and Welfare} 
  \label{tab:outcomes} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
& \multicolumn{4}{c}{\textbf{Council Size and Welfare -- Main Estimates}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Infant}} & 
 \multicolumn{1}{c}{\small{Postneonatal}} &
  \multicolumn{1}{c}{\small{School}} & 
  \multicolumn{1}{c}{\small{Education}} \\ 
   & \multicolumn{1}{c}{\small{Mortality}} & 
 \multicolumn{1}{c}{\small{Mortality}} &
  \multicolumn{1}{c}{\small{Enrollment}} & 
  \multicolumn{1}{c}{\small{Quality}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & -2.01$^{**}$ & -0.90$^{*}$ & 2.52$^{***}$ & -0.03 \\ 
  & (0.79) & (0.48) & (0.80) & (0.13) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{12299} & \multicolumn{1}{c}{5441} & \multicolumn{1}{c}{10156} & \multicolumn{1}{c}{8597} \\ 
N Right & \multicolumn{1}{c}{1029} & \multicolumn{1}{c}{672} & \multicolumn{1}{c}{686} & \multicolumn{1}{c}{669} \\ 
Eff N Left & \multicolumn{1}{c}{501} & \multicolumn{1}{c}{324} & \multicolumn{1}{c}{206} & \multicolumn{1}{c}{409} \\ 
Eff N Right & \multicolumn{1}{c}{381} & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{291} \\ 
BW Loc Poly & 7.527 & 7.564 & 5.425 & 9.05 \\ 
BW Bias & 11.95 & 11.01 & 10.58 & 15.107 \\ [1.0ex] 
\midrule 
& \multicolumn{4}{c}{\textbf{Council Size and Welfare -- Placebo Estimates}} \\ \cline{2-5}
& \multicolumn{1}{c}{\small{(5)}} 
 & \multicolumn{1}{c}{\small{(6)}} 
 & \multicolumn{1}{c}{\small{(7)}} 
 & \multicolumn{1}{c}{\small{(8)}} \\
& \multicolumn{1}{c}{\small{Infant}} & 
 \multicolumn{1}{c}{\small{Postneonatal}} &
  \multicolumn{1}{c}{\small{School}} & 
  \multicolumn{1}{c}{\small{Education}} \\ 
   & \multicolumn{1}{c}{\small{Mortality}} & 
 \multicolumn{1}{c}{\small{Mortality}} &
  \multicolumn{1}{c}{\small{Enrollment}} & 
  \multicolumn{1}{c}{\small{Quality}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & -1.14 & -0.19 & 0.44 & -0.10 \\ 
  & (1.05) & (0.22) & (0.61) & (0.11) \\ 
    \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{10613} & \multicolumn{1}{c}{17236} & \multicolumn{1}{c}{9040} & \multicolumn{1}{c}{7528} \\ 
N Right & \multicolumn{1}{c}{2715} & \multicolumn{1}{c}{3624} & \multicolumn{1}{c}{1802} & \multicolumn{1}{c}{1738} \\ 
Eff N Left & \multicolumn{1}{c}{1210} & \multicolumn{1}{c}{780} & \multicolumn{1}{c}{1210} & \multicolumn{1}{c}{1255} \\ 
Eff N Right & \multicolumn{1}{c}{890} & \multicolumn{1}{c}{664} & \multicolumn{1}{c}{728} & \multicolumn{1}{c}{752} \\ 
BW Loc Poly & 4.838 & 2.807 & 6.364 & 6.873 \\ 
BW Bias & 7.143 & 8.041 & 9.68 & 10.291 \\
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

Table \ref{tab:outcomes} shows that increasing the city council by one legislator generates welfare improvement in municipalities. In health care, increasing the legislature size decreases infant mortality in 2.01 deaths per 1,000 children born. This impact is significant in absolute terms and represents an effect of 0.14 in standard deviations. Moreover, postneonatal mortality decreases by 0.90 casualties per 1,000 births, which remains around the 0.14 standard deviations improvement as infant mortality. Infant mortality is a primary outcome for welfare, and it correlates with several other indicators of well-being \citep{world2005health}. The fact that infant mortality improved in the Brazilian case shows that the increased mayor's support mechanism results in sizable welfare effects. As expected, placebo regressions were statistically insignificant.

@@

>>>>

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: 
  - Umberto Mignozzeti[^umberto]
  - Gabriel Cepaluni[^gabriel]
  - Danilo Freire[^danilo]
date: \today
abstract: "How does legislature size impact public service provision? Despite the importance of institutional design for democratic governance, the effect of particular legislative features on citizen welfare remains little understood. In this paper, we use a formal model to show that increasing legislature size improves public goods provision. We argue that changes in bargaining costs depend on whether additional legislators share the Executive's party affiliation: More opposition members reduce the Executive's public goods provision, while more government-aligned members increase it. We test this theory by exploiting sharp discontinuities in city council size in Brazil. We show that an additional city councilor has a 91% chance of belonging to the mayoral coalition, and this significantly improves primary school enrollment and infant mortality measures. To explore possible mechanisms, we surveyed former councilors and analyzed 346,553 bills proposed between 2005 and 2008. This paper has implications for the design of representative institutions."
abstractspacing: double
keywords: Legislature Size; Legislative Bargaining; Local Public Goods Provision; Regression Discontinuity Design; Brazil
jelcodes: H41; H80; I31; I38
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

[^umberto]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>. Corresponding author.

[^gabriel]: Associate Professor, Department of International Relations, São Paulo State University, <gabi.cepal@gmail.com>.

[^danilo]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <dfreire@lincoln.ac.uk>, <https://danilofreire.github.io>.

@@

Third, our article also contributes to a growing literature on municipal partisan politics. Most papers in this area estimate the effects of the mayor's party on political outcomes. For instance, \citet{de2016mayoral} and \citet{gerber2011mayors} find differences in governance when comparing Democratic and Republican mayors. Specifically about Brazil, \citet{gouvea2021partisanship} shows that social policy expenditure increase with left-wing mayors. \citet{frey2020larger}, using discontinuities in city council size after the 2009 constitutional amendment that reverted the thresholds we study here, shows that changes in council size lower the support for candidates backed by the municipal mayors. Our paper reinforces the view that partisanship matters locally, even in countries with weak ideological linkage.

Finally, this paper contributes to a growing literature on the causes and effects of legislature sizes. Most papers on legislature size study the trade-off between representation and efficiency \citep{frederick2008people, jacobs2015explaining, muzzio1989size, Stigler1976}. However, some recent articles explore the effects of legislature size on lawmaking and partisanship. \citet{kirkland2014chamber} shows that larger legislatures make it harder for politicians to know each other and collaborate, making them rely more heavily on party cues and committee structures. In considering the causes of changes in legislature sizes, \citet{jacobs2015explaining} show that population shifts and economic crises cause enlargement and shrinkage, respectively. \citet{marland2019fewer} questions the economic crises mechanism, showing that political elites shrink legislatures mostly to signal to the public the priority of their austerity agenda. In this article, we show that legislature size affects welfare by changing the partisan composition of legislatures.

@@

The design of institutions is crucial for long term welfare \citep{acemoglu2012nations, buenodemesquita2005logic, north1989constitutions} and democratic consolidation \citep{gandhi2019institutional, lijphart2018institutional}. One particular problem that each polity faces is determination of the size of its legislative bodies. The literature is extensive, and modern attempts to provide a solution to the legislature size problem started as early as in the eighteenth century, with the Federalists versus Anti-Federalists debates \citep{muzzio1989size, Stigler1976}.

The legislature size literature highlights that the number of legislators causes tension between representation and efficiency \citep{frederick2008people, jacobs2015explaining, muzzio1989size, rush2013conclusion, Stigler1976, taagepera1972size}. Larger legislatures improve representation as the election thresholds decrease and the ratio of population-per-politician decreases. This facilitates the electoral success of ethnic and political minorities \citep{allenrepresentation}. Conversely, the improvements in representation are countered by the decline in efficiency in larger legislatures. The enlarged size makes it harder for members to coordinate and collaborate without partisanship and legislative structures \citep{kirkland2014chamber, rush2013conclusion}. It also increases the potential number of veto players, decreasing the policymaking efficiency \citep{tsebelis2002veto}. Balancing these two incentives is crucial for the optimal legislative composition \citep{auriol2012optimal, taagepera1972size}.

However, the legislature size scholarship has overlooked two questions: first, does legislature size matter in contexts where the legislative bodies are not the pivotal decision-maker? And second, what is the ultimate effect of legislature size on welfare? On the one hand, legislature size's effects on welfare represent a significant gap in the political economy literature. It is likely generated by the predominance of the "law of 1/n" scholarship, which emphasized larger legislatures' overspending consequence. Since the seminal paper by \citet{Weingastetal1981}, 30 papers have systematically tested the effects of legislature size on government spending \citep{freire2021legislature}. Of these papers, only \citet{lewis2019legislature} investigates its effects on welfare, showing that larger legislatures have a negative welfare effect.\footnote{\citet{koppel2004public} also shows, using a repeated game where the proposal prerogative rotates in a legislature, that larger legislatures increase the chances of non-cooperative outcomes, lowering public goods provision.} \citet{frey2020larger} analyses the Brazilian case after the 2009 legislature size change and finds a null effect in service provision from 2012 to 2016. These results suggest that the legislature size and welfare nexus needs a dedicated theoretical and empirical assessment.

Despite the division of power, the Executive dominates the debates on public service provision in most mayor-council (or presidential) systems. For instance, while the New York Times cited the New York City Mayor Bill De Blasio 2,084 times in 2021, it mentioned the 49 seating city councilors 440 times in the same period. To avoid the COVID crisis when the Executive was in the spotlight, we expanded the search for the past five years. In this new search, Bill De Blasio was mentioned 8,978 while the 49 seating city councilors were mentioned 2,046 times.\footnote{These figures were computed using the Dow Jones Factiva Database in November 2021.} The mayor was cited, on average, over four times more than all the city councilors combined. This Executive dominance has implications for policymaking, as coalitions and affinity with the Executive are going to be crucial for service provision \citep{buenodemesquita2005logic}.

@@

The solution concept we use in this model is the subgame perfect Nash equilibrium. This game requires that the strategies follow a Nash equilibrium in each subgame. We find the solution using backward induction. In the model with infinitely repeated proposals, any strategy can be shown to be a Nash equilibrium. Thus, a natural assumption is to extend the equilibrium concept to require the equilibrium to be stationary. A stationary subgame perfect equilibrium requires that, at each given point of the game, if a politician accepts an offer at time $k+1$, she should take the same offer at time $k$ \citep{rubinstein1982perfect}. Stationarity gives us a method to find a proposal that would be accepted at any stage of the game. Hence, we can characterize a sequence of offers for each point $k$ in time. For the mayor, the optimal offer is at $k=0$, representing a no delay in policy implementation. We assume that players have no dominated strategies, even when there is insufficient support for the policy.

@@

\begin{table}[!htpb] \centering 
  \caption{Political Effects of Increasing the Number of Legislators} 
  \label{tab:bargain} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
& \multicolumn{4}{c}{\textbf{Panel A: Representation, Competition, and Legislation Approval}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Num. Female}} & 
 \multicolumn{1}{c}{\small{Num. Non-white}} &
  \multicolumn{1}{c}{\small{Candidates}} & 
  \multicolumn{1}{c}{\small{Prop. Laws}} \\ 
   & \multicolumn{1}{c}{\small{councillors}} & 
 \multicolumn{1}{c}{\small{councillors}} &
  \multicolumn{1}{c}{\small{Per Seat}} & 
  \multicolumn{1}{c}{\small{Approved Council}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & 0.22 & 0.60 & -0.19 & -0.02 \\ 
  & (0.38) & (0.85) & (1.11) & (0.08) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5183} & \multicolumn{1}{c}{239} & \multicolumn{1}{c}{5184} & \multicolumn{1}{c}{3424} \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{158} & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{269} \\ 
Eff N Left & \multicolumn{1}{c}{195} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{179} & \multicolumn{1}{c}{227} \\ 
Eff N Right & \multicolumn{1}{c}{145} & \multicolumn{1}{c}{47} & \multicolumn{1}{c}{132} & \multicolumn{1}{c}{141} \\ 
BW Loc Poly & 8.623 & 2.958 & 8.144 & 10.976 \\ 
BW Bias & 13.565 & 4.922 & 12.018 & 16.708 \\[1.0ex] 
\midrule 
& \multicolumn{4}{c}{\textbf{Panel B: Partisan Alignment and Bureaucratic Appointments}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(5)}} 
 & \multicolumn{1}{c}{\small{(6)}} & & \\
 & \multicolumn{1}{c}{\small{Mayoral}} & 
 \multicolumn{1}{c}{\small{Num. Politically}} & & \\
 & \multicolumn{1}{c}{\small{Coalition Size}} & 
 \multicolumn{1}{c}{\small{Bureaucrats}} & 
  & \\[1ex] \cline{2-2} \cline{3-3}
 LATE & 0.91$^{*}$ & 104.49$^{*}$ & & \\ 
  & (0.50) & (61.53) & & \\ 
    \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{5179} & \multicolumn{1}{c}{15536} & & \\ 
N Right & \multicolumn{1}{c}{343} & \multicolumn{1}{c}{1027} & & \\ 
Eff N Left & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{372} & & \\ 
Eff N Right & \multicolumn{1}{c}{162} & \multicolumn{1}{c}{343} & & \\ 
BW Loc Poly & 10.025 & 6.242 & & \\ 
BW Bias & 15.889 & 10.243 & & \\ 
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}

@@

\begin{figure}[H]
  \begin{center}
  \caption{Legislation Approval Patterns in the Municipalities Closer to the City Council Size Thresholds}
    \begin{minipage}{1 \linewidth}
      \centerline{\includegraphics[width=6in,angle=0,height=6.5in]{../figures/lawsApproved.pdf}}
      \bigskip
      \label{LEGAPPR}
    \end{minipage}
  \end{center}
\end{figure}

@@

Table \ref{tab:outcomes} shows that increasing the city council by one legislator generates welfare improvement in municipalities. In health care, increasing the legislature size decreases infant mortality in 2.01 deaths per 1,000 children born. This impact is significant in absolute terms and represents an effect of 0.14 in standard deviations. Moreover, postneonatal mortality decreases by 0.90 casualties per 1,000 births, which remains around the 0.14 standard deviations improvement as infant mortality. Infant mortality is a primary outcome for welfare, and it correlates with several other indicators of well-being \citep{world2005health}. The fact that infant mortality improved in the Brazilian case shows that the increased mayor's support mechanism results in large welfare effects. As expected, placebo regressions were statistically insignificant.

@@

\begin{table}[!htpb] \centering 
  \caption{Legislature Size and Welfare} 
  \label{tab:outcomes} 
\scalebox{0.9}{
\begin{threeparttable}
\begin{tabularx}{\textwidth}{@{\extracolsep{\fill}}l c c c c }
\toprule
& \multicolumn{4}{c}{\textbf{Council Size and Welfare -- Main Estimates}} \\ \cline{2-5}
 & \multicolumn{1}{c}{\small{(1)}} 
 & \multicolumn{1}{c}{\small{(2)}} 
 & \multicolumn{1}{c}{\small{(3)}} 
 & \multicolumn{1}{c}{\small{(4)}} \\
 & \multicolumn{1}{c}{\small{Infant}} & 
 \multicolumn{1}{c}{\small{Postneonatal}} &
  \multicolumn{1}{c}{\small{School}} & 
  \multicolumn{1}{c}{\small{Education}} \\ 
   & \multicolumn{1}{c}{\small{Mortality}} & 
 \multicolumn{1}{c}{\small{Mortality}} &
  \multicolumn{1}{c}{\small{Enrollment}} & 
  \multicolumn{1}{c}{\small{Quality}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & -2.01$^{**}$ & -0.90$^{*}$ & 2.52$^{***}$ & -0.03 \\ 
  & (0.79) & (0.48) & (0.80) & (0.13) \\ 
  \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{12299} & \multicolumn{1}{c}{5441} & \multicolumn{1}{c}{10156} & \multicolumn{1}{c}{8597} \\ 
N Right & \multicolumn{1}{c}{1029} & \multicolumn{1}{c}{672} & \multicolumn{1}{c}{686} & \multicolumn{1}{c}{669} \\ 
Eff N Left & \multicolumn{1}{c}{501} & \multicolumn{1}{c}{324} & \multicolumn{1}{c}{206} & \multicolumn{1}{c}{409} \\ 
Eff N Right & \multicolumn{1}{c}{381} & \multicolumn{1}{c}{244} & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{291} \\ 
BW Loc Poly & 7.527 & 7.564 & 5.425 & 9.05 \\ 
BW Bias & 11.95 & 11.01 & 10.58 & 15.107 \\ [1.0ex] 
\midrule 
& \multicolumn{4}{c}{\textbf{Council Size and Welfare -- Placebo Estimates}} \\ \cline{2-5}
& \multicolumn{1}{c}{\small{(5)}} 
 & \multicolumn{1}{c}{\small{(6)}} 
 & \multicolumn{1}{c}{\small{(7)}} 
 & \multicolumn{1}{c}{\small{(8)}} \\
& \multicolumn{1}{c}{\small{Infant}} & 
 \multicolumn{1}{c}{\small{Postneonatal}} &
  \multicolumn{1}{c}{\small{School}} & 
  \multicolumn{1}{c}{\small{Education}} \\ 
   & \multicolumn{1}{c}{\small{Mortality}} & 
 \multicolumn{1}{c}{\small{Mortality}} &
  \multicolumn{1}{c}{\small{Enrollment}} & 
  \multicolumn{1}{c}{\small{Quality}} \\[1ex] \cline{2-2} \cline{3-3} \cline{4-4} \cline{5-5}
LATE & -1.14 & -0.19 & 0.44 & -0.10 \\ 
  & (1.05) & (0.22) & (0.61) & (0.11) \\ 
    \hdashline \\[-1.8ex] 
 N Left & \multicolumn{1}{c}{10613} & \multicolumn{1}{c}{17236} & \multicolumn{1}{c}{9040} & \multicolumn{1}{c}{7528} \\ 
N Right & \multicolumn{1}{c}{2715} & \multicolumn{1}{c}{3624} & \multicolumn{1}{c}{1802} & \multicolumn{1}{c}{1738} \\ 
Eff N Left & \multicolumn{1}{c}{1210} & \multicolumn{1}{c}{780} & \multicolumn{1}{c}{1210} & \multicolumn{1}{c}{1255} \\ 
Eff N Right & \multicolumn{1}{c}{890} & \multicolumn{1}{c}{664} & \multicolumn{1}{c}{728} & \multicolumn{1}{c}{752} \\ 
BW Loc Poly & 4.838 & 2.807 & 6.364 & 6.873 \\ 
BW Bias & 7.143 & 8.041 & 9.68 & 10.291 \\
\bottomrule
\end{tabularx}
\begin{tablenotes}
\small
\item \textbf{Note:} $^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1. RD local linear estimates using \citet{CCT2014} optimal bandwidth quadratic selection and triangular kernel. Robust standard errors, clustered at the municipal level, in parentheses. Controls: population; GDP per capita; number of seats in 2000; year; and dummy for northeast region. \textit{N Left} and \textit{N Right} represent the total number of observations on the left and right sides of the thresholds (untreated). \textit{Eff N Left} and \textit{Eff N Right} are the number of cases within the bandwidth. \textit{BW Loc Poly} is the Bandwidth used to compute the Local Average Treatment Effect (LATE). \textit{BW Bias} is the Bandwidth used to compute the standard errors.
\end{tablenotes}
\end{threeparttable}}
\end{table}
